Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
getwd()

pf<- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot( x= age, y= friend_count,data = pf)

#alternative code:
#qplot(age,friend_count,data=pf)
```

***

#### What are some things that you notice right away?
Response:
people under 30 have a lot of friends. 
people over 90 also seem to have a lot but that's most likely incorrect info. 
***

### ggplot Syntax
Notes:
always use aes wrapper

```{r ggplot Syntax}
ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point() + xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13,90)

```

#### What do you notice in the plot?
Response:
majority of lies below the age of 25.

peak around 69

***

### Coord_trans()
Notes:
# if using jitter:

'''
ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13,90)+
  coord_trans( y = "sqrt")

```{r Coord_trans()}
ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13,90)+
  coord_trans( y = "sqrt")

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point(alpha = 1/20) + 
  xlim(13,90)+
  coord_trans( y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:
jitter can add postive and negative noise to each of our points
```{r Alpha and Jitter}

ggplot(aes(x= age, y= friend_count),data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h=0)) + 
  xlim(13,90)+
  coord_trans( y = "sqrt")


```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:
n = n() is only used with the summarize function and reports how many people really are in ecah group

we need to pass a dataframe or a grouping at the beginning of each function; we save that into a new variable and pass this new variable into the next function.

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

age_group <- group_by(pf, age) #passing df/grouping to a var
pf.fc_by_age <- summarise(age_group,  
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) #passing the var to next function

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Create your plot!

following code produces an error:
```{r}
library(dplyr)
pf.fc_by_ages<- pf %>%
  group_by(age) %>%
  summarise(age_group,  
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(age,20)
```
  
  arrange(age)
#this code also produces the same results as above so pass it to a variable. 

```{r Conditional Means Plot}

ggplot(aes(x= age, y= friend_count_mean),data = pf.fc_by_age) +
  geom_line() 
  
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y = friend_count), data = pf)+
  xlim(13,90) +
  geom_point(alpha = 0.05,
             position = position_jitter(h =0),
             color = 'orange') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list( probs = 0.1),
           linetype = 1, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = 0.5), color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = 0.9),
           linetype = 2, color = 'blue')

```

#### What are some of your observations of the plot?
Response:
spike at 69 and higher friend count for people under 30. all consistent with previous observations. 

but friend count falls between ages 30 to 50
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
age = pf$age
count = pf$friend_count
cor(age,count)
```
```{r}
#ALTERNATIVE CODE

cor.test(pf$age, pf$friend_count, method = 'pearson')

#can also do the following:

with(pf, cor.test(age,friend_count, method = 'pearson'))
```
Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***-0.027

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))

#adding method will also yield the same
#pearson method is default with cor.test
#can add mehtod and change mehtod too
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
names(pf)
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(0,quantile(pf$www_likes_received, 0.95))+
  ylim(0,quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')


```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.
strong postive correlation

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:
strong correlation. 
***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:
Type str(functionName) to find the order of the parameters and learn more about the parameters of an R function

following is the way to load mitchell.
itscontained in the alr3 package.

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data("Mitchell")
?Mitchell
```

Create your plot!

```{r Temp vs Month}
names(Mitchell)
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()

#qplot syntax
qplot(data = Mitchell, Month, Temp)

str(qplot)
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:
range of the months is 203 so we choose the lower and upper limit in the break and 12 is for 12 months
```{r Making Sense of Data}
#range(Mitchell$Month)
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()+
  scale_x_discrete(breaks = seq(0,203,12))
```

***

### A New Perspective
The cor and cor.test functions determine the strength of a linear relationship, but they may miss other relationships in the data.
There are other measures of associations that can detect this. The dcor.ttest() function in the energy package implements a non-parametric test of the independence of two variables. While the Mitchell soil dataset is too coarse to identify a significant dependency between "Month" and "Temp", we can see the difference between dcor.ttest and cor.test through other examples, like the following:

x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)

##
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
  
What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x= age, y= friend_count_mean),data = pf.fc_by_age) +
  geom_line() 
head(pf.fc_by_age)
pf.fc_by_age[17:19, ]
```

***

### Age with Months Means
age_group <- group_by(pf, age) #passing df/grouping to a var
pf.fc_by_age <- summarise(age_group,  
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) #passing the var to next function
```{r Age with Months Means}
pf$age_with_months <- pf$age + (12- pf$dob_month) / 12

pf.fc_by_agemonths <- summarise(age_group,
                      friend_count_mean = mean(friend_count),
                     friend_count_median = median(friend_count),
                       n = n())     

arrange(age_with_months)

```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_agemonths <- summarise(age_group,
                      friend_count_mean = mean(friend_count),
                     friend_count_median = median(friend_count),
                       n = n())     

arrange(pf.fc_by_agemonths, age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

